---
title: "HW2"
author: "Sai Akhil Rayapudi"
date: "2023-02-07"
output: pdf_document
---

**Problem 1**\
```{r warning=FALSE, message=FALSE}
library(readr)
library(stringr)
library(tidyverse)
dataf <- 
read_csv("/Users/saiakhil/Desktop/DS Coursework/Sem 1/IDMP/HW2/googleplaystore.csv",
         show_col_types = FALSE)
#data reading
str(dataf)
unique(dataf$Installs) #review data
which(dataf$Installs == 0) #finding rows in the data frame with installs as zero
dataf[9149, ]

df1 <- dataf[-9149,]
#dropping this row(9149) as the number of installs is 0 and it doesn't have any relevant 
#data in the other columns too.

head(df1)
df1[9149,] 
#reviewing the data in this specific row

df2 <- df1[,!names(df1) %in% c("Last Updated","Current Ver","Android Ver")]
#Dropping columns Last Updated, Current Ver and Android Ver as they don't provide any 
#meaningful correlation to the user's install trends.

str(df2)

df3 <- df2[,-10]
#Category and Genres columns give the same data, so dropping Genres column to make
#the data frame simple.

summary(df3$Rating)
x <- df3[is.na(df3$Rating), ] #finding null values

library(dplyr)
table(x$Category)
#One interesting observation from here is that every app in the entertainment 
#category has a rating which is not the case with every other category in the data

y <- df3 %>% 
  group_by(Category) %>% 
  mutate_at(vars(Rating), 
            ~replace_na(., 
                        mean(., na.rm = TRUE)))
#Replaced null values in the column Rating based on the mean of the values of 
#its own category

unique(y$Rating)
z <- ungroup(y)

z$Rating <- round(z$Rating ,digit=1)
#Rounding off mean to 1 digit after the decimal point to get proper data

```
**Observations:**\
Citing data source- Google Play Store Apps. (2019, February 3). Kaggle. Retrieved February 12, 2023, from https://www.kaggle.com/datasets/lava18/google-play-store-apps \
I have taken this data from kaggle, it describes the data of 10,000 playstore apps \

Variables: \
App - App Name \
Category - Category of the app \
Rating - Rating of the app shared by user \
Reviews - Number of reviews shared by the user \
Size - Total size of the App \
Installs - Number of times the app was installed by the users \
Type - Type of the app i.e., Paid/Free \
Price - Price of the app if it is of Paid type \
Content Rating - Viewer discretion rating of the App content

Preprocessing steps: \
1. Data cleaning by removing null values \
2. Removing rows that are contributing no important insights \
3. Finding and dropping features that are irrelevant \
4. Imputing values with mean to make the data consistent

**Problem 2**\
```{r fig.width=15, warning=FALSE, message=FALSE}
#Graph 1 Category vs Installs

z$Installs <- gsub(",","",z$Installs)
#Removing commas in the data by replacing with a blank space
unique(z$Installs)

library(stringr)
z$Installs <- str_sub(z$Installs, end = -2)
unique(z$Installs)
#Removing the '+' symbol at the end to make the data more meaningful for plots

class(z$Installs)
z$Installs <- as.numeric(z$Installs)
#Converting the column from string to numeric to plot the data
class(z$Installs)


cat_installs <- z %>% group_by(Category) %>% 
  summarise(sum_installs=sum(Installs),
            .groups = 'drop') %>% as.data.frame()

#Plot Category vs Installs
ggplot(data=cat_installs, aes(x=Category, y=sum_installs)) +
  geom_bar(stat="identity")+
  geom_text(aes(label = sum_installs), vjust = -0.2,size=2.25)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Plot between category and the number of app installs in that category")

```

**Observations:**\
Apps in the 'GAME' category has the most number of installs, suggesting high preference among the users \
Similarly apps in the 'EVENTS' has the least number of installs suggesting lower user base

```{r fig.width=15, warning=FALSE, message=FALSE}
#Graph 2 Avg Rating based on Category
df_new <- z
df_new

data_new1 <- df_new %>% group_by(Category) %>% 
  summarise(mean_rating=mean(Rating),
            .groups = 'drop') %>%
  as.data.frame()
#Grouping the 'Rating' column based on the 'Category' column

str(data_new1)
#data_new1 is the data frame containing average rating based on category

#Plot mean_Rating vs Category
ggplot(data=data_new1, aes(x=Category, y=mean_rating)) +
   geom_point(color="blue")+
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
   ggtitle("Plot between Category and the mean of rating in each category")

data_new2 <- df_new %>% group_by(Category) %>% 
  as.data.frame()
#data grouped by 'Category' column
str(data_new2)

#Plot Rating vs Category
ggplot(data = data_new2, mapping = aes(x=Category, y=Rating)) +
geom_boxplot()+
theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
ggtitle("Box Plot between Category and rating in each category")

```

**Observations:**\
From the above graph, we can observe that rating is high for the apps in the 'Events' Category. \
An interesting correlation from the first 2 graphs is that 'Events' is the one with least number of installs but the users who have used them have rated them highly

```{r fig.width=15, warning=FALSE, message=FALSE}
#Graph3 Total App installs based on category for paid apps
data_new3 <- data_new2
data_new3 <- data_new3[data_new3$Type == "Paid", ]
str(data_new3)
#Filtering based on the value of Type column which is Paid

paid_installs <- data_new3 %>% group_by(Category) %>% 
  summarise(new_installs=sum(Installs),
            .groups = 'drop') %>% as.data.frame()
#Grouping the data by category and finding the sum of installs for each category

#Plot Category vs Installs for paid apps
ggplot(data=paid_installs, aes(x=Category, y=new_installs)) +
  geom_bar(stat="identity")+
  geom_text(aes(label = new_installs), vjust = -0.2,size=2.5)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Plot between Category and app installs in each category for paid apps")
```

**Observations:**\
From the above graph we can see that most of the paid apps installed belong to the 'Family' category. \
On a contrast, paid apps of 'Events' category has been installed only once.

**Problem 3**\
```{r}
library(readr)
dataf <- 
read_tsv("/Users/saiakhil/Desktop/DS Coursework/Sem 1/IDMP/HW2/NCAA-D1-APR-2003-14/DS0001/26801-0001-Data.tsv"
         ,show_col_types = FALSE)
#loading data frame
library(dplyr) #loading dplyr library for select function
dataff <- select(dataf,SCL_UNITID:ACADEMIC_YEAR,starts_with("APR_RATE_")) #selecting columns required
str(dataff)
library("reshape2") #loading reshape library for the melt function
dataff1 <- subset(dataff,select= -ACADEMIC_YEAR) 
#dropping ACADEMIC YEAR column to rearrange the data set
colnames(dataff1) <-c("SCL_UNITID","SCL_NAME","SPORT_CODE","SPORT_NAME","2014",
        "2013","2012","2011","2010","2009","2008","2007","2006","2005","2004")
#changing the column names of data set for the melt function to work properly 
dataff2 <- melt(dataff1,id.vars=c("SCL_UNITID","SCL_NAME","SPORT_CODE","SPORT_NAME"),
                variable.name = "YEAR", value.name="APR")
#melt function to transform the data frame to wide to long
dataff2<-dataff2[!(dataff2$APR<0),]
#Removing rows containing APR's with negative values
str(dataff2)

#Plot year vs APR
library(ggplot2)
ggplot(data = dataff2, mapping = aes(x=YEAR, y=APR)) +
  geom_boxplot()+
  coord_cartesian(ylim = c(550,1000))+
  ggtitle("Trend of APR's over the given time frame of 2004-2014 ")
```
**Observations:**\
From the graph, we can see that the APR values follow an increasing trend from the year 2004 to 2014.

**Problem 4**\
```{r fig.width=12, warning=FALSE, message=FALSE}
dataff3 <- dataff2[dataff2$SPORT_CODE != 38, ]
#Removing SPORT_CODE of 38 as it is a mised gender sport.
# Men 1-18 and Women 19-37

unique(dataff3$SPORT_CODE)

dataff3$GENDER <- ifelse (dataff3$SPORT_CODE %in% c(1:18),"Men","Women")
#Adding a column called Gender and assigning values based on the SPORT code
str(dataff3)

#Plot of year vs APR split out by GENDER
library(ggplot2)
ggplot(data = dataff3, mapping = aes(x=YEAR, y=APR,color=GENDER)) +
  geom_boxplot()+
  facet_grid(~GENDER)+
  coord_cartesian(ylim = c(550,1000))+
  ggtitle("Trend of APR's over the given time frame of 2004-2014 broken out by gender")
```
**Observations:**\
We can see that the APR's of Women are on a higher side compared with the APR's of Men for the same year. \
APR's in general follow an increasing trend when observed over the timeframe from 2004 to 2014. 

**Problem 5**\
```{r fig.width=10,fig.height=10, warning=FALSE, message=FALSE}

pattern <- "Men's." 
replacement <- ""
dataff3$SPORT_NAME <- gsub(pattern,replacement,dataff3$SPORT_NAME)
str(dataff3)
#Removing Men's prefix from the sport name
unique(dataff3$YEAR)

pattern1 <- "Women's." 
dataff3$SPORT_NAME <- gsub(pattern1,replacement,dataff3$SPORT_NAME)
str(dataff3)
#Removing Women's prefix from the sport name

#Plot APR vs SPORT split by gender
library(ggplot2)
ggplot(data = dataff3, mapping = aes(x=APR, y=SPORT_NAME,color=GENDER)) +
  geom_boxplot()+
  coord_cartesian(xlim = c(550,1000))+
  ggtitle("Plot between APR and SPORT_NAME split by gender")
```

**Observations:**\
After observing the distribution of APR's for Men's and Women's teams for each sport, \
we can see that APR's are same for Men and Women in below sports namely: \
Tennis, Skiing, Gymnastics, Golf, Fencing, Cross Country \
Apart from this, APR's are close for Men and Women in Volleyball and Ice Hockey